---
title: "hw05_ninadh"
author: "Ninadh"
date: "October 19, 2018"
output: html_document
---


```{r}
library(gapminder)
library(tidyverse)
library(scales)
library(ggplot2)
library(plotly)
library(knitr)
```

Part 1: Factor management

- Reorder the level of country
- Drop Oceania

```{r}
summary(gapminder)
```

A better structure of the gapminder dataset

```{r}
gapminder %>%
  group_by(continent) %>%
  knitr::kable() %>%
  head()
```

```{r}
gapminder %>%
summarize(
n_row = nrow(gapminder),
cont_levels = nlevels(gapminder$country),
coun_levels = nlevels(gapminder$continent)) %>%
knitr::kable(col.names = c("Rows", "Countries", "Continents"))
```

- Reordering the level of country and gdpPercap

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  group_by(country) %>%
  select(country, gdpPercap) %>%
  knitr::kable(col.names = c("Country", "GDP Percap")) %>%
  head(10)
```

Using the "arrange" function

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  select(country, gdpPercap, year) %>%
  mutate(max_GDP = max(gdpPercap)) %>%
  arrange(max_GDP) %>%
  knitr::kable() %>%
  head(20)
```

This shows the max gdpPercap by an Asian country in a certain year.

visualisation using ggplot

```{r}
arranged_gap <- gapminder %>%
  filter(continent == "Asia") %>%
  select(country, gdpPercap, year) %>%
  mutate(max_GDP = max(gdpPercap)) %>%
  arrange(max_GDP)
```

```{r}
arranged_gap %>%
  ggplot(aes(max_GDP, year, colour = country)) +
  geom_point()
```

this is weird. according to this graph, Yemen had the max gdp over the years.

```{r}
arranged_gap %>%
  filter(country == "Bangladesh") %>%
  ggplot(aes(max_GDP, year)) +
  geom_point()
```

ok. dont know whats going on with GDP. Let me try with lifeExp.

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  group_by(country) %>%
  select(country, lifeExp) %>%
  knitr::kable(col.names = c("Country", "Life Expectancy")) %>%
  head(10)
```

```{r}
arranged_gap <- gapminder %>%
  filter(continent == "Asia") %>%
  select(country, lifeExp, year) %>%
  mutate(max_LEx = max(lifeExp)) %>%
  arrange(max_LEx)
```


```{r}
arranged_gap %>%
  ggplot(aes(lifeExp, year, colour = country)) +
  geom_point()
```

