---
title: "hw05_ninadh"
author: "Ninadh"
date: "October 19, 2018"
output: github_document
---


```{r}
library(gapminder)
library(tidyverse)
library(scales)
library(ggplot2)
library(plotly)
library(knitr)
```

Part 1: Factor management

- Reorder the level of country
- Drop Oceania

```{r}
summary(gapminder)
```

A better structure of the gapminder dataset

```{r}
gapminder %>%
  group_by(continent) %>%
  knitr::kable() %>%
  head()
```

```{r}
gapminder %>%
summarize(
n_row = nrow(gapminder),
cont_levels = nlevels(gapminder$country),
coun_levels = nlevels(gapminder$continent)) %>%
knitr::kable(col.names = c("Rows", "Countries", "Continents"))
```

- Reordering the level of country and gdpPercap

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  group_by(country) %>%
  select(country, gdpPercap) %>%
  knitr::kable(col.names = c("Country", "GDP Percap")) %>%
  head(10)
```

Using the "arrange" function

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  select(country, gdpPercap, year) %>%
  mutate(max_GDP = max(gdpPercap)) %>%
  arrange(max_GDP) %>%
  knitr::kable() %>%
  head(20)
```

This shows the max gdpPercap by an Asian country in a certain year.

visualisation using ggplot

```{r}
arranged_gap <- gapminder %>%
  filter(continent == "Asia") %>%
  select(country, gdpPercap, year) %>%
  mutate(max_GDP = max(gdpPercap)) %>%
  arrange(max_GDP)
```

```{r}
arranged_gap %>%
  ggplot(aes(max_GDP, year, colour = country)) +
  geom_point()
```

this is weird. according to this graph, Yemen had the max gdp over the years.

```{r}
arranged_gap %>%
  filter(country == "Bangladesh") %>%
  ggplot(aes(max_GDP, year)) +
  geom_point()
```

ok. dont know whats going on with GDP. Let me try with lifeExp.

```{r}
gapminder %>%
  filter(continent == "Asia") %>%
  group_by(country) %>%
  select(country, lifeExp) %>%
  knitr::kable(col.names = c("Country", "Life Expectancy")) %>%
  head(10)
```

```{r}
nlevels(gapminder$lifeExp)
nlevels(gapminder$gdpPercap)
nlevels(gapminder$pop) 
nlevels(gapminder$country)  
nlevels(gapminder$continent)
nlevels(gapminder$year)
```



```{r}
arranged_gap <- gapminder %>%
  filter(continent == "Asia") %>%
  select(country, lifeExp, year) %>%
  mutate(max_LEx = max(lifeExp)) %>%
  arrange(max_LEx)
```

```{r}
gapminder %>%
  filter(year > 1987) %>%
  filter(continent == "Asia") %>%
  select(country, year, lifeExp)
```


```{r}
arranged_lifeEx <- gapminder %>%
  filter(year > 1987) %>%
  filter(continent == "Asia") %>%
  select(country, year, lifeExp)
```


```{r}
arranged_lifeEx %>%
  ggplot(aes(lifeExp, year, colour = country)) +
  geom_point()
```

Dropping Oceania now.

```{r}
gapminder %>%
  filter(continent != "Oceania") %>%
  droplevels()
```

```{r}
levels(gapminder$continent)
```



Checking for the continents after dropping Oceania.

```{r}
noOc_gap <- gapminder %>%
  filter(continent != "Oceania") %>%
  droplevels()
```

```{r}
levels(noOc_gap$continent)
```

Plotting for lifeExp after dropping Oceania

```{r}
noOc_gap %>%
  ggplot(aes(continent, lifeExp, colour = continent)) +
  geom_jitter() +
  labs (x = "Continent",
        y = "Life Expectancy",
        title = "Life expectancy after dropping Oceania") +
  theme_bw()
```

Part 2: File input and output (File I/O)

- Have been using gapminder all along and will stick to it.
- Will use the filtered factor levels created in Part 1.

```{r}
arranged_lifeEx <- gapminder %>%
  filter(year > 1987) %>%
  filter(continent == "Asia") %>%
  select(country, year, lifeExp)
```


Trying to save the file as csv. as instructed in the assigment.

```{r}
write.csv(arranged_lifeEx, file = "STAT545_hw05")
```


Trying to import using read.csv as instructed

```{r}
read_csv("STAT545_hw05")
```


Part 3: Visualisation design

- going to use plotly for this

```{r}
library(plotly)
```


```{r}
noOc_gap %>%
  ggplot(aes(lifeExp, year)) +
  geom_point(aes(colour = pop)) +
  facet_wrap(~ continent) +
     theme_bw() +
  labs(x= "Life Expectancy", y = "Year") +
    theme(axis.text = element_text(size = 8),
          axis.title = element_text(size = 20),
          strip.background = element_rect(fill = "purple"))
```

this graph shows that Asia had very high population (light blue) at some points, and its position with respect to Year and Life expectance.

```{r}
plot_ly(noOc_gap, 
        x = ~lifeExp, 
        y = ~year, 
        z = ~pop,
        type = "scatter",
        mode = "markers",
        opacity = 0.5) 
```

I would like to try the 3D version.

```{r}
plot_ly(noOc_gap, 
        x = ~lifeExp, 
        y = ~year, 
        z = ~pop,
        type = "scatter3d",
        mode = "markers",
        opacity = 0.5) 

# "3D" did not work. try "3d". case sensitive. work eith extending at the edges to enlarge or shrink with trackpad.
```


Part 4: Writing figures to file

```{r}
hw05_fig <- noOc_gap %>%
  ggplot(aes(lifeExp, year)) +
  geom_point(aes(colour = pop)) +
  facet_wrap(~ continent) +
     theme_bw() +
  labs(x= "Life Expectancy", y = "Year") +
    theme(axis.text = element_text(size = 8),
          axis.title = element_text(size = 20),
          strip.background = element_rect(fill = "purple"))
```


```{r}
ggsave("hw05_plot.png", hw05_fig)
```

changing the scaling of the figure

```{r}
ggsave("hw05_plot2.png", hw05_fig, width = 50, height = 40, units = "cm", dpi = 600)
```

hw05_plot is better looking than hw05_plot2

